\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{parskip}

\title{Game Project: Skateboard Adventure}
\author{Ryo Nakayama \\ Student number: 250151349}
\date{\today}

\begin{document}

\maketitle

\section*{Innovative Features and Technical Implementations}

This skateboard adventure game incorporates several innovative features that enhance gameplay experience and demonstrate advanced programming concepts.

\textbf{Dynamic Camera System:} Implemented a sophisticated camera tracking system that smoothly follows the character's horizontal movement. The camera uses world coordinates and screen translation, creating an immersive side-scrolling experience that keeps the player centered while revealing the expansive game world.

\textbf{Advanced Physics Engine:} Developed a comprehensive physics system including gravity simulation, jump mechanics with realistic arc trajectories, and multi-surface collision detection. The character can land on platforms, fall into canyons, and interact with various environmental elements using precise collision radius calculations.

\textbf{Enemy Characters:} Created Detective Conan-inspired criminal silhouettes with simple patrol movement. Enemies move back and forth within defined boundaries, changing direction at limits, and feature detailed sprite rendering with glowing eyes and shadowed forms, adding atmospheric tension to gameplay.

\textbf{Multi-State Game Management:} Implemented robust state management handling multiple game conditions simultaneously: normal play, falling into voids, platform interactions, enemy encounters, level completion, and game over scenarios. Each state triggers appropriate visual feedback and game logic responses.

\textbf{Visual Polish and Detail:} Enhanced visual appeal through detailed sprite art including skateboard components (deck, trucks, wheels with highlights), environmental elements (oak and pine trees, layered mountains, animated clouds), and atmospheric color palettes that create depth and immersion.

\textbf{Debug System:} Integrated a comprehensive debug overlay (activated with 'Z' key) displaying real-time game state information, collision data, and entity positions, facilitating development and demonstrating systematic programming approach.

These innovations demonstrate mastery of object-oriented programming, game state management, and creative problem-solving in interactive media development.

\end{document}
